Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?

### 2016-03-24 article/article.Rmd
+ article fixes + remove randomly generated prioritizations from surrogacy analysis

To understand the relationship between the surrogate-based targets used to generate a prioritization and the proportion of genetic variation it secures, we generated a collection of single-species prioritizations. For each species, we first generated two sets of prioritizations using increasing environmental and geographic-based targets (separately). Prioritizations were generated using surrogate-based targets ranging from `r (min(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target)*100)` % and `r (max(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target)*100)` % in increments of `r diff(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target[1:2])*100` % ($n = `r rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.optimal.replicates`$ per target). To determine if these surrogate-based prioritizations performed better then random, we then generated another two sets of prioritizations by randomly selecting planning units that were occupied by the species ($n = `r rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.random.replicates`$ per surrogate target). To account for variation in the number of selected planning units in the target-based prioritizations, the number of selected planning units in the randomly generated prioritizations was equal to the target-based prioritizations. Randomly generated prioritizations associated with negative values--because they secured only a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis.

To determine if the use of surrogate-based targets can actually improve prioritizations in real-world conservation planning scenarios, we investigated the effectiveness of surrogate-based targets using three scenarios with increasing levels of complexity. To assess the benefits of surrogate-based targets under the simplest of circumstances, our first scenario involved generating single-species prioritizations for each species. However, most large-scale prioritizations involve a comprehensive set of species. By forcing solutions to secure individuals in multiple communities, this may result in prioritizations that secure a greater proportion of intra-specific genetic variation for each species. To test for this effect, our second scenario involved generating multi-species prioritizations. Finally, most large-scale prioritizations also explicitly consider opportunity cost to ensure prioritizations are cost-effective. Our third scenario involved generating multi-species prioritizations that satisfied targets for minimal opportunity cost.

For each scenario, we generated three types of prioritizations using (1) only amount-based targets, (2) both amount- and surrogate-based targets, and (3) both amount- and genetic-based targets. The purely amount-based prioritizations served as our baseline--representing prioritizations typically generated in conservation planning exercises. We used `r (rapr.params.LST[[MODE]][['single.species']]$amount.target*100)` % amount-based targets in all prioritizations to ensure an adequate proportion of habitat was secured for each species. We set surrogate- and genetic-based targets as `r (rapr.params.LST[[MODE]][['single.species']]$amount.target*100)` % to ensure that most of the respective variation was secured for each species. To accomodate variation in the high number of optimal solutions for the single-species amount-based prioritizations, we generated `r (rapr.params.LST[[MODE]][['single.species']]$amount.replicates)` replicate prioritizations. Since all other contexts were likely to be associated with few optimal solutions, we generated only `r rapr.params.LST[[MODE]][['multi.species']]$genetic.replicates` replicates.

## Statistical analysis
We investigated the effectiveness of prioritizations generated using surrogate-based targets using generalized linear mixed-effects models [GLMMs; using a logit link and the $\texttt{lme4}$ R package; @r469]. To determine  if the environmental and geographic variables were effective surrogates, we fit a full model to the proportion of genetic variation that the first collection of prioritizations. This full model contained categorical predictor variables to indicate the surrogate-based targets used to generate them, the type of genetic variation (adaptive or neutral), the method used to generate the prioritization (optimal or random), and the interactions between these variables. To determine if surrogate-based targets improved the effectiveness of prioritizations under various scenarios, we fit a full model to the proportion of genetic variation that second collection of prioritizations secured. This second full model contained categorical predictor variables to indicate the scenario (single-species, multi-species, multi-species with cost), the combination of targets used to generate the prioritizations (amount-based targets, amount and surrogate-based targets, amount and genetic-based targets), the type of genetic variation measured (adaptive or neutral), and the interactions between them. All model contained intercept-level random effects to accommodate variation between species. Both full models were subject to a step-wise term deletion routine using $\chi^{2}$ tests to assess significance (terms were retained when $P < 0.05$). Finally, we conducted post-hoc analysis on each full model using Tukey contrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].

## Results
<!-- Surrogacy analysis -->
Prioritizations generated using higher environmental and geographic surrogate-based targets were found to secure greater proportions of adaptive and neutral genetic variation (Figure 1). The proportion of genetic variation that these prioritizations secured was substantially more than the genetic variation secured by randomly generated prioritizations (${\chi^{2}}_{`r NA`,`r NA`} = `r  NA`$, $P < `r NA`$). Generally, prioritizations needed to capture at least `r NA` % of the environmental and geographic surrogates to secure 90 % of the species' genetic variation. Additionally, we detected differences in the effectiveness of the environmental and geographic surrogates (${\chi^{2}}_{`r NA`,`r NA`} = `r  NA`$, $P < `r NA`$). Increases in environmental surrogate-based targets resulted prioritizations that secured more adaptive genetic variation, than geographic surrogate-based targets for neutral variation. These results demonstrate that environmental and geographic variables are effective surrogates for genetic variation for these plant species.

<!-- Scenario analysis -->
The addition of environmental and geographic surrogate-based targets to amount-based targets was not found to substantially improve the effectiveness of prioritizations under any of the scenarios (Figure 2). No significant differences were detected between prioritizations generated using different combinations of targets (${\chi^{2}}_{`r NA`,`r NA`} = `r  NA`$, $P < `r NA`$), or under any of the different scenarios (${\chi^{2}}_{`r NA`,`r NA`} = `r NA`$, $P < `r NA`$). The only difference was that the inclusion of surrogate-based targets tended to result in prioritizations that secured slightly more adaptive than neutral genetic variation. Under the simplest of circumstances, the single-species amount-based prioritizations secured most of the species' adaptive (mean `r NA` $\pm$ `r NA`) and neutral (mean `r NA` $\pm$ `r NA`) genetic variation. These proportions were not significantly different to those associated with prioritizations generated using additional surrogate-based targets ($Z = `r NA`$, $P < `r NA`$) or genetic-based targets ($Z = `r NA`$, $P < `r NA`$). We found similar patterns in the multi-species scenario ${\chi^{2}}_{`r NA`,`r NA`} = `r NA`$, $P < `r NA`$) and the multi-species with opportunity cost scenario ${\chi^{2}}_{`r NA`,`r NA`} = `r NA`$, $P < `r NA`$). Although the prioritizations had similar levels of performance, there were differences in their spatial distributions (Figure 3). In particular, the inclusion of cost data in the multi-species prioritizations had the effect of concentrating conservation effort towards areas with higher elevation. Overall, these results suggest that while environmental and geographic surrogate variables are effective surrogates, they are not needed to secure most of the intra-specific genetic variation.

## Discussion
We aimed to investigate the effectiveness of environmental and geographic surrogates for genetic variation in conservation planning. Our study is the first to test this in a realistic context. We found that environmental and geographic variables are effective surrogates for genetic variation in conservation planning. But we also found that just using amount-based targets--typically used in conservation planning--result in prioritizations that secure most of the adaptive and neutral intra-specific genetic variation. Our results support existing guidelines on using a comprehensive set of species in conservation planning.

>>>>

To determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation, we generated a collection of single-species prioritizations. For each species, we first generated two sets of prioritizations using increasing environmental and geographic-based targets (separately). Prioritizations were generated using surrogate-based targets ranging from `r (min(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target)*100)` % and `r (max(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target)*100)` % in increments of `r diff(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target[1:2])*100` % ($n = `r rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.optimal.replicates`$ per target). Prioritizations associated with negative values--because they secured only a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis.

To determine if the use of surrogate-based targets can actually improve prioritizations in real-world conservation planning scenarios, we investigated the effectiveness of surrogate-based targets using three scenarios with increasing levels of complexity. To assess the benefits of surrogate-based targets under the simplest of circumstances, our first scenario involved generating single-species prioritizations for each species. However, most large-scale prioritizations involve a comprehensive set of species. By forcing solutions to secure individuals in multiple communities, this may result in prioritizations that secure a greater proportion of intra-specific genetic variation for each species. To test for this effect, our second scenario involved generating multi-species prioritizations. Finally, most large-scale prioritizations also explicitly consider opportunity cost to ensure that prioritizations are cost-effective. Our third scenario involved generating multi-species prioritizations that satisfied targets for minimal opportunity cost.

For each scenario, we generated three types of prioritizations using (1) only amount-based targets, (2) both amount- and surrogate-based targets, and (3) both amount- and genetic-based targets. The purely amount-based prioritizations served as our baseline--representing the prioritizations typically generated in conservation planning exercises. We used `r (rapr.params.LST[[MODE]][['single.species']]$amount.target*100)` % amount-based targets in all prioritizations to ensure an adequate proportion of habitat was secured for each species. We set surrogate- and genetic-based targets as `r (rapr.params.LST[[MODE]][['single.species']]$amount.target*100)` % to ensure that most of the respective variation was secured for each species. To accomodate variation in the high number of optimal solutions for the single-species amount-based prioritizations, we generated `r (rapr.params.LST[[MODE]][['single.species']]$amount.replicates)` replicate prioritizations. Since all other contexts were likely to be associated with few optimal solutions, we generated only `r rapr.params.LST[[MODE]][['multi.species']]$genetic.replicates` replicates.

## Statistical analysis
We investigated the effectiveness of prioritizations generated using surrogate-based targets using generalized linear mixed-effects models [GLMMs; using a logit link and the $\texttt{lme4}$ R package; @r469]. To determine  if the environmental and geographic variables were effective surrogates, we fit a full model to the proportion of genetic variation that the first collection of prioritizations. This full model contained categorical predictor variables to indicate the surrogate-based targets used to generate them, the type of genetic variation (adaptive or neutral), and the interactions between these variables. To determine if surrogate-based targets improved the effectiveness of prioritizations under various scenarios, we fit a full model to the proportion of genetic variation that second collection of prioritizations secured. This second full model contained categorical predictor variables to indicate the scenario (single-species, multi-species, multi-species with cost), the combination of targets used to generate the prioritizations (amount-based targets, amount and surrogate-based targets, amount and genetic-based targets), the type of genetic variation measured (adaptive or neutral), and the interactions between them. All models contained intercept-level random effects to accommodate variation between species. We conducted a post-hoc analysis on each full model using Tukey contrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].

## Results
<!-- Surrogacy analysis -->
The environmental and geographic-based variables were effective surrogates for adaptive and genetic variation (Figure 1). The prioritizations generated using greater surrogate-based targets secured more genetic variation (${\chi^{2}}_{`r NA`,`r NA`} = `r  NA`$, $P < `r NA`$). The environmental-based surrogates tended to secure a substantially higher proportion of genetic variation than the geographic-based prioritizations (${\chi^{2}}_{`r NA`,`r NA`} = `r  NA`$, $P < `r NA`$). Only particularly high geographic-based targets ($\geq 0.95 %$) were able to secure a moderate proportion of neutral variation. Overall, there appears to be a non-linear relationship between the surrogate-based target used to generate a prioritization and the proportion of genetic variation secured.

<!-- Scenario analysis -->
To determine if the addition of environmental and geographic surrogate-based targets could improve the effectiveness of prioritizations, we generated prioritizations according to three different scenarios (Figure 2). The single-species prioritizations scenarios showed similar selection frequencies across the study area, regardless of whether they were generated using amount-only targets, amount- and surrogate-based targets, or amount- and genetic-based targets. On the other hand, the multi-species prioritizations showed different selection frequencies depending on the targets used to generate them. The addition of surrogate- or genetic-based targets resulted in prioritizations that contained substantially more planning units. Furthermore, the inclusion of cost data in the multi-species prioritizations resulted in even more planning units being selected. These prioritizations also tended to concentrate conservation effort towards areas with higher elevation. The inclusion of surrogate- and genetic- based targets tended to result in prioritizations that captured more genetic variation.

The addition of environmental and geographic surrogate-based targets had the greatest effect on the proportion of neutral variation secured in a prioritization. In both the single species and multi-species scenarios, the inclusion of surrogate- and genetic-based targets resulted in prioritizations that secured substantially more neutral variation. However, this effect was not present in the multi-species with cost scenario. Unlike neutral variation, the addition of environmental and geographic surrogate-based targets had little effect on the proportion of adaptive variation secured in a prioritization. Although the inclusion of surrogate- and genetic-based targets resulted in more adaptive variation being secured in the single species scenarios, this effect was not present in either the multi-species or the multi-species with cost scenarios. This is likely due to the fact that the multi-species prioritizations were constrained to secure each species in a variety of different communities, and therefore, potentially different habitats.

## Discussion
We aimed to investigate the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation in conservation planning. Our study is the first to test this in a multi-species context. Overall, we found that environmental and geographic variables are not necessarily effective surrogates for genetic variation in conservation planning--though they can be useful to for identifying suitable amount-based targets. Additionally, we found that surrogate-based targets can supplement amount-based targets and improve the proportion of neutral variation secured in a prioritization if the amount-based targets are too low. Our results suggest that in most cases--barring extremely unusual spatial patterns of intra-specific genetic variation--conservation planners can secure sufficient adaptive and neutral variation by following conventional guidelines.

### 2016-03-24 article/figures.Rmd
+ article fixes + remove randomly generated prioritizations from surrogacy analysis

```{r, echo=FALSE, message=FALSE, fig.width=6, fig.height=4.5, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations. Panel (a) shows the relationship between environmental surrogate targets used to generate a prioritization and the proportion of adaptive genetic variation it secures. Similarly, panel (b) shows that for geographic surrogate targets and neutral genetic variation. Box-plots show inter-specific variation in the proportion of genetic variation secured using different surrogate-based targets. Letters indicate significant differences ($P < 0.05$).'}
# prepare data
env.correlation.DF$Surrogate.target <- factor(env.correlation.DF$Surrogate.target)
env.correlation.DF2 <- env.correlation.DF %>%
	group_by(Surrogate.target, Method, Type) %>% 
	summarize(genetic.held=max(genetic.held, na.rm=TRUE)+0.1) %>%
	ungroup() %>% mutate(Combination=as.character(paste0(Surrogate.target,'.',Method,'.',Type))) %>% 
	left_join(
		data.frame(
			letter=unname(toupper(cld(posthoc.correlation.GLHT)$mcletters$Letters)),
			Combination=as.character(names(toupper(cld(posthoc.correlation.GLHT)$mcletters$Letters))),
			stringsAsFactors=FALSE
		),
		by='Combination'
	) %>% data.frame()

geo.correlation.DF$Surrogate.target <- factor(geo.correlation.DF$Surrogate.target)
geo.correlation.DF2 <- geo.correlation.DF %>%
	group_by(Surrogate.target, Method, Type) %>% 
	summarize(genetic.held=max(genetic.held, na.rm=TRUE)+0.1) %>%
	ungroup() %>% mutate(Combination=as.character(paste0(Surrogate.target,'.',Method,'.',Type))) %>% 
	left_join(
		data.frame(
			letter=unname(toupper(cld(posthoc.correlation.GLHT)$mcletters$Letters)),
			Combination=as.character(names(toupper(cld(posthoc.correlation.GLHT)$mcletters$Letters))),
			stringsAsFactors=FALSE
		),
		by='Combination'
	) %>% data.frame()

# make figure
p1 <- ggplot(aes(x=Surrogate.target,y=genetic.held, color=Method),
		data=env.correlation.DF) +
	geom_boxplot(position=position_dodge(0.9)) +
	xlab('Environmental surrogate target (%)') + 
	ylab('Adaptive genetic\nvariation secured (%)') +
	geom_text(aes(x=0.55, y=1.125, label='(a)'), color='black', size=5) +
	theme_classic() +
	theme(legend.position=c(0.9,0.4)) +
	scale_color_manual(name='Prioritization', values=c('Optimal'='black', 'Random'='grey50')) +
	geom_text(aes(x=Surrogate.target, y=genetic.held, color=Method, label=letter),
		data=env.correlation.DF2, position=position_dodge(0.9))

p2 <- ggplot(aes(x=Surrogate.target,y=genetic.held, color=Method),
			data=geo.correlation.DF) +
	geom_boxplot(position=position_dodge(0.9)) +
	xlab('Geographic surrogate target (%)') + 
	ylab('Neutral genetic\nvariation secured (%)') +
	geom_text(aes(x=0.55, y=1.125, label='(b)'), color='black', size=5) +
	theme_classic() +
	theme(legend.position='none') +
	scale_color_manual(name='Prioritization', values=c('Optimal'='black', 'Random'='grey50')) +
	geom_text(aes(x=Surrogate.target, y=genetic.held, color=Method, label=letter),
		data=geo.correlation.DF2, position=position_dodge(0.9))

>>>>

```{r, echo=FALSE, message=FALSE, fig.width=6, fig.height=4.5, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations. Panel (a) shows the relationship between environmental surrogate targets used to generate a prioritization and the proportion of adaptive genetic variation it secures. Similarly, panel (b) shows that for geographic surrogate targets and neutral genetic variation. Box-plots show inter-specific variation in the proportion of genetic variation secured using different surrogate-based targets. Letters indicate significant differences ($P < 0.05$).'}
# prepare data
env.correlation.DF$Surrogate.target <- factor(env.correlation.DF$Surrogate.target)
env.correlation.DF <- env.correlation.DF %>%
	group_by(Surrogate.target, Type, Species) %>% 
	summarize(genetic.held=mean(genetic.held, na.rm=TRUE)) %>%
	data.frame()

env.correlation.DF2 <- env.correlation.DF %>%
	group_by(Surrogate.target, Type) %>% 
	summarize(genetic.held=max(genetic.held, na.rm=TRUE)+0.1) %>%
	ungroup() %>% mutate(Combination=as.character(paste0(Surrogate.target,'.',Type))) %>% 
	left_join(
		data.frame(
			letter=unname(toupper(cld(posthoc.correlation.GLHT)$mcletters$Letters)),
			Combination=as.character(names(toupper(cld(posthoc.correlation.GLHT)$mcletters$Letters))),
			stringsAsFactors=FALSE
		),
		by='Combination'
	) %>% data.frame()

geo.correlation.DF$Surrogate.target <- factor(geo.correlation.DF$Surrogate.target)
geo.correlation.DF <- geo.correlation.DF %>%
	group_by(Surrogate.target, Type, Species) %>% 
	summarize(genetic.held=mean(genetic.held, na.rm=TRUE)) %>%
	data.frame()

geo.correlation.DF2 <- geo.correlation.DF %>%
	group_by(Surrogate.target, Type) %>% 
	summarize(genetic.held=max(genetic.held, na.rm=TRUE)+0.1) %>%
	ungroup() %>% mutate(Combination=as.character(paste0(Surrogate.target,'.',Type))) %>% 
	left_join(
		data.frame(
			letter=unname(toupper(cld(posthoc.correlation.GLHT)$mcletters$Letters)),
			Combination=as.character(names(toupper(cld(posthoc.correlation.GLHT)$mcletters$Letters))),
			stringsAsFactors=FALSE
		),
		by='Combination'
	) %>% data.frame()

# make figure
p1 <- ggplot(aes(x=Surrogate.target,y=genetic.held),
		data=env.correlation.DF) +
	geom_boxplot(position=position_dodge(0.9)) +
	xlab('Environmental surrogate target (%)') + 
	ylab('Adaptive genetic\nvariation secured (%)') +
	geom_text(aes(x=0.55, y=1.125, label='(a)'), color='black', size=5) +
	theme_classic() +
	theme(legend.position='none') +
	geom_text(aes(x=Surrogate.target, y=genetic.held, label=letter),
		data=env.correlation.DF2, position=position_dodge(0.9))

p2 <- ggplot(aes(x=Surrogate.target,y=genetic.held),
			data=geo.correlation.DF) +
	geom_boxplot(position=position_dodge(0.9)) +
	xlab('Geographic surrogate target (%)') + 
	ylab('Neutral genetic\nvariation secured (%)') +
	geom_text(aes(x=0.55, y=1.125, label='(b)'), color='black', size=5) +
	theme_classic() +
	theme(legend.position='none') +
	geom_text(aes(x=Surrogate.target, y=genetic.held, label=letter),
		data=geo.correlation.DF2, position=position_dodge(0.9))
